---
title: "Getting & Cleaning Quiz 3"
author: "tedscott"
date: "January 21, 2016"
output: html_document
---

Quiz 3

Q1: The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf

Create a logical vector that identifies the households on greater than 10 acres who sold more than $10,000 worth of agriculture products. Assign that logical vector to the variable agricultureLogical. Apply the which() function like this to identify the rows of the data frame where the logical vector is TRUE.

which(agricultureLogical)

What are the first 3 values that result?

```{r}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(fileURL, destfile="./data/IdahoHousing2006.csv")
idaho <- read.csv("./data/IdahoHousing2006.csv")
idahoDF <- tbl_df(idaho)
idahoDF

# from code book, variables ACR, where ACR = 3 have 10 or more acres
# variable AGS = 6 means sold more than $10k
# create a new column agriculturalLogical
idahoDF <- mutate(idahoDF, agriculturalLogical=(ACR==3 & AGS==6))
idahoDF
head(idahoDF$agriculturalLogical)
which(idahoDF$agriculturalLogical)
#idahoDF %>% which(agriculturalLogical)

```
Q2:
Using the jpeg package read in the following picture of your instructor into R

https://d396qusza40orc.cloudfront.net/getdata%2Fjeff.jpg

Use the parameter native=TRUE. What are the 30th and 80th quantiles of the resulting data? (some Linux systems may produce an answer 638 different for the 30th quantile)

```{r}
install.packages("jpeg")
library(jpeg)
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fjeff.jpg"
download.file(fileURL, destfile="./data/jeffimage.jpg", mode="wb")#, mode="wb")

img <- readJPEG("./data/jeffimage.jpg", native=TRUE)
# from http://stackoverflow.com/questions/15283447/how-to-read-jpeg-in-r-2-15
quantile(img, probs=c(0.3,0.8))


```

Q3:
Load the Gross Domestic Product data for the 190 ranked countries in this data set:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv

Load the educational data from this data set:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv

Match the data based on the country shortcode. How many of the IDs match? Sort the data frame in descending order by GDP rank (so United States is last). What is the 13th country in the resulting data frame?

Original data sources:

http://data.worldbank.org/data-catalog/GDP-ranking-table

http://data.worldbank.org/data-catalog/ed-stats

```{r}
file1 <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
file2 <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
download.file(file1, destfile="./data/GDP-ranking-table.csv")
download.file(file2, destfile="./data/ed-stats.csv")
GDP <- read.csv("./data/GDP-ranking-table.csv", stringsAsFactors = FALSE)
ED <- read.csv("./data/ed-stats.csv", stringsAsFactors = FALSE)
names(GDP)
names(ED)
head(GDP,5)
tail(GDP,5)
head(ED)

# the GDP data is in rows 5 - 194, column X is CountryCode, Gross.domestic.product.2012 is
# the ranking and X.2 is the country name, X.3 is the GDP in millions $
# so it is time to subset it
GDPsub <- GDP[5:194,c("X","Gross.domestic.product.2012","X.2","X.3")]
names(GDPsub) <- c("CountryCode","GDPRank","CountryName","GDP_millions")

# merge based on CountryCode in GDP and CountryCode in ED
mergedData <- merge(x = GDPsub, y=ED, by.x = "CountryCode", by.y = "CountryCode", all=TRUE)
mergedDF <- tbl_df(mergedData)
names(mergedDF)

# count NAs to see how many merged (=190)
nrow(mergedDF) - sum(is.na(mergedDF$GDPRank))

head(arrange(mergedDF, desc(GDPRank)),13)

## NOTE - apparently 190 is wrong! Must be 189?

# seems that I need to convert the GDPRank to be numeric
mergedDF$GDPRank <- as.numeric(mergedDF$GDPRank)
head(arrange(mergedDF, desc(GDPRank)),13)

# 13th ranked is St Kitts

```


Q4:
What is the average GDP ranking for the "High income: OECD" and "High income: nonOECD" group?

```{r}

# subsets
mergedDF %>% filter(Income.Group=="High income: OECD") %>% summarise(avgGDPRank=mean(GDPRank,na.rm=T))

mergedDF %>% filter(Income.Group=="High income: nonOECD") %>% summarise(avgGDPRank=mean(GDPRank,na.rm=T))



```

Q5:
Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group. How many countries

are Lower middle income but among the 38 nations with highest GDP?


```{r}

# cut into 5 quantiles based on GDPRank
mergedDF2 <- mutate(mergedDF, GDPRankGroup=cut2(GDPRank, g=5))
table(mergedDF2$GDPRankGroup, mergedDF2$Income.Group)

# based on table, looks like 5 countries


```

