---
title: "Getting and Cleaning Quiz 1"
author: "tedscott"
date: "January 8, 2016"
output: html_document
---

Quiz 1

Question 1 - Idaho property values
How many properties are worth $1,000,000 or more?

```{r}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(fileURL, destfile="./IdahoHousing.csv")

idahoHousing <- read.csv("IdahoHousing.csv")
table(idahoHousing$VAL)

```

Question 2 - no code, FES has more than one variable in the column

Question 3 - working with XLSX for Nat Gas
Read rows 18-23 and columns 7-15 into R and assign the result to a variable called dat
What is the value of sum(dat$Zip*dat$Ext,na.rm=T)

```{r}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
download.file(fileURL, destfile="./getdata-data-DATA.gov_NGAP.xlsx",method="curl")
library(xlsx)
NAGP <- read.xlsx("./getdata-data-DATA.gov_NGAP.xlsx", sheetIndex = 1, header=TRUE)
head(NAGP)

# can read specific rows and columns
colIndex <- 7:15
rowIndex <- 18:23
dat <- read.xlsx("./getdata-data-DATA.gov_NGAP.xlsx", sheetIndex = 1, 
                              colIndex=colIndex, rowIndex = rowIndex)
sum(dat$Zip*dat$Ext,na.rm=T)

```

Question 4
Read the XML data on Baltimore restaurants from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml

How many restaurants have zipcode 21231?

```{r}
library(XML)
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
# above doesn't work due to https
# read file in with chrome and saved it locally
filename <- "./restaurants.xml"
doc <- xmlTreeParse(filename, useInternal=T)
rootNode <- xmlRoot(doc)
xmlName(rootNode)
names(rootNode)
#rootNode[[1]]
# get all the zipcodes
zipcodes <- xpathSApply(rootNode, "//zipcode",xmlValue)
table(zipcodes)

```

Question 5

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv

using the fread() command load the data into an R object called: DT
The following are ways to calculate the average value of the variable: pwgtp15
broken down by sex. Using the data.table package, which will deliver the fastest user time?

DT[,mean(pwgtp15),by=SEX]

sapply(split(DT$pwgtp15,DT$SEX),mean)

mean(DT$pwgtp15,by=DT$SEX)

tapply(DT$pwgtp15,DT$SEX,mean)

rowMeans(DT)[DT$SEX==1]; rowMeans(DT)[DT$SEX==2]

mean(DT[DT$SEX==1,]$pwgtp15); mean(DT[DT$SEX==2,]$pwgtp15)

```{r}
library(data.table)


# already have the csv
filename <- "acs.csv"
DT <- fread(filename)

system.time(DT[,mean(pwgtp15),by=SEX])

# single runs don' yield much info, so using replicate to get avgs
system.time(replicate(500,{DT[,mean(pwgtp15),by=SEX]}))
system.time(replicate(500,{sapply(split(DT$pwgtp15,DT$SEX),mean)}))
system.time(replicate(500,{mean(DT$pwgtp15,by=DT$SEX)}))
system.time(replicate(500,{tapply(DT$pwgtp15,DT$SEX,mean)}))
system.time(replicate(500,{rowMeans(DT)[DT$SEX==1]}))
system.time(replicate(500,{mean(DT[DT$SEX==1,]$pwgtp15)}))


```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
