---
title: "ReadingFromMySQL"
author: "tedscott"
date: "December 19, 2015"
output: html_document
---

MySQL

Root account: root
password: Pa$$word1

user account: default
password: Co2SiO4!

need to install RTools and RMySQL packages

this page is helpful: http://www.ahschulz.de/2013/07/23/installing-rmysql-under-windows/

MySQL is in C:\Program Files\MySQL\MySQL Server 5.7

```{r}
library(RMySQL)

# connect to local MySQL DB using built in table
con  <- dbConnect(MySQL(), host="127.0.0.1", port= 3306, user="default",password = "Co2SiO4!", dbname="sakila")

# see what DBs there are
result <- dbGetQuery(con, "show databases;")
result

# connect to UCSC Genome web mysql db
# see http://genome.ucsc.edu/goldenPath/help/mysql.html
ucscDB <- dbConnect(MySQL(), user="genome", host="genome-mysql.cse.ucsc.edu")
result <- dbGetQuery(ucscDB, "show databases;")
result
dbDisconnect(ucscDB)

# now use human genome 19 db on ucsc genome db
hg19 <- dbConnect(MySQL(), user="genome", db="hg19", host="genome-mysql.cse.ucsc.edu")
allTables <- dbListTables(hg19)
length(allTables)
allTables[1:5]

# get a count from a specific table
# the fields are the columns
dbListFields(hg19, "affyU133Plus2")
dbGetQuery(hg19, "select count(*) from affyU133Plus2")

```

Now let's put contents from the table into a data frame

```{r}
# read from the table and put contents into a DF
affyData <- dbReadTable(hg19,"affyU133Plus2")
head(affyData)

# select a subset of the data
query <- dbSendQuery(hg19, "select * from affyU133Plus2 where misMatches between 1 and 3")
affyMis <- fetch(query)
quantile(affyMis$misMatches)
hist(affyMis$misMatches)

# select only a few rows
affyMisSmall <- fetch(query,n=10)

# clear the query from the server
dbClearResult(query)
dim(affyMisSmall)

# disconnect
dbDisconnect(hg19)

# can also do on.exit(dbDisconnect(hg19)) to disconnect when R is closed
# check out http://www.r-bloggers.com/mysql-and-r/ for more fun stuff

```




You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
